Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-07-28 Documents/Paper/ms.tex
A word changed
Paragraphs talking about sinewaves results

When using DNS,
the proposals typically should be very heavy tailed. When
DNS is exploring a distribution close to the prior, large proposals will
generally be needed to explore the prior efficiently.
When DNS is exploring a distribution constrained to
very high values of the likelihood function, much smaller proposals will be
appropriate. Rather than attempting expensive tuning (since the number of
distributions involved may number in the hundreds or even thousands), we will
apply the heavy tailed proposal distributions and simply accept that there will
be some waste. If a parameter $u$ has a uniform prior between 0 and 1, then
a heavy-tailed proposal is defined by:
\begin{eqnarray}
u' &=& \textnormal{mod}\left(u + 10^{1.5 - 6a}b, 1\right)
\end{eqnarray}
where $a \sim \textnormal{Uniform}(0, 1)$ and $b\sim \mathcal{N}(0,1)$.
This is like a standard gaussian/normal random walk proposal, but with a
non-fixed step size, resulting in a scale mixture of normals.
When
$a$ is close to zero, the size of the proposal is of order 10, which (because
of the mod function) effectively randomises $u'$ from the prior. If $a$ is
close to 1, the scale of the proposal is roughly $10^{-4}$ times the prior
width. The coefficient of $a$ was set to 6 because smaller proposals are
usually not necessary. The value 1.5 in the exponent was found to optimize the
performance of this proposal when sampling from $U(0, 1)$ and $\mathcal{N}(0,1)$
priors.

@@

>>>>

When using DNS,
the proposals typically should be very heavy tailed. When
DNS is exploring a distribution close to the prior, large proposals will
generally be needed to explore the prior efficiently.
When DNS is exploring a distribution constrained to
very high values of the likelihood function, much smaller proposals will be
appropriate. Rather than attempting expensive tuning (since the number of
distributions involved may number in the hundreds or even thousands), we will
apply the heavy tailed proposal distributions and simply accept that there will
be some waste. If a parameter $u$ has a uniform prior between 0 and 1, then
a particular choice of heavy-tailed proposal is defined by:
\begin{eqnarray}
u' &=& \textnormal{mod}\left(u + 10^{1.5 - 6a}b, 1\right)
\end{eqnarray}
where $a \sim \textnormal{Uniform}(0, 1)$ and $b\sim \mathcal{N}(0,1)$.
This is like a standard gaussian/normal random walk proposal, but with a
non-fixed step size, resulting in a scale mixture of normals.
When
$a$ is close to zero, the size of the proposal is of order 10, which (because
of the mod function) effectively randomises $u'$ from the prior. If $a$ is
close to 1, the scale of the proposal is roughly $10^{-4}$ times the prior
width. The coefficient of $a$ was set to 6 because smaller proposals are
usually not necessary. The value 1.5 in the exponent was found to optimize the
performance of this proposal when sampling from $U(0, 1)$ and $\mathcal{N}(0,1)$
priors.

@@

\subsection{Results}
We ran DNS on the simulated data shown in Figure~\ref{fig:sinewave_data}.
The shape of the likelihood function, with respect to the prior mass (the
usual likelihood curve associated with Nested Sampling), us shown in
Figure~\ref{fig:sinewaves_likelihood}. Two phase transitions can be seen. The
first, at $\log(X) \approx -10$ nats,
separates ``noise-only'' models (where the
entire dataset is accounted for by the noise term) from ``one-sinusoid'' models
where $N=1$. At $\log(X) \approx -28$ nats, a second phase transition separates
$N=1$ models from $N=2$ models.

If we were to do this analysis by trying to explore the posterior directly,
rather than by using Nested Sampling, it would be difficult to jump between
the $N=1$ solution and the $N=2$ solution, because the posterior distribution
would be composed of a mixture of a small-volume but high-likelihood spike
and a large-volume but (relatively) low likelihood
slab. Note that this phase transition still exists if we condition on $N=2$,
so is not caused by the trans-dimensional model.
